EoS-insensitive relations after GW170817

### 2019-01-06 paper/bin-love-after-gw170817.tex
C-Love universality

\title{Universal relations after GW170817\kc{Here and elsewhere: I suggest we use the term "EoS-insensitive" instead of "universal" relations}}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state) of \kc{cold} supranuclear matter, accessible primarily in neutron stars\kc{this statement makes nuclear experiment people not happy}, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately equation of state insensitive ``binary Love universal relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of equation of state independent relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" universal relations.
% What universal relations do
Such universal relations\kc{so far we have said "EoS-insensitive", "EoS-independent", and "universal". We should pick a name and stick with it. I suggest "EoS-insensitive"} allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing the uncertainty of the dominant tidal effect in parameter estimation\kc{You mean $\tilde{\Lambda}$? In our paper with Carl and Aaron we showed that the binary Love relation doesn't affect the $\tilde{\Lambda}$ measurement much.}.
% What we will do
\kc{Add a sentence saying that even the more precise relations have some EoS variation which can cause a systematic error which can be marginalized over}
In this document, we explain how one can reduce equation of state variation for both binary Love and I-Love-Q universal relations, which helps to reduce systematic errors on the tidal measurement of binary neutron star mergers\kc{but we show that the systematic effects of these relations is very small until next-G detectors.}. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing systematic errors on tidal measurement due to the equation of state variation with statistical errors and comment on whether one can safely use these universal relations with future gravitational wave observations.\kc{We should give the answer to this here: Yes!}
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
%Intro
The dependence of pressure on density for \kc{cold} supranuclear matter found primarily inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS up to the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy}, making NSs ideal laboratories for constraining ultra dense nuclear matter\kc{heavy ion collisions supposedly go to higher densities, but they are T-dependent}.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the mass-radius relationship of NSs have been used to place limits on NSs\kc{limits on what?}~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications \kc{potentially} suffer from large systematic errors due to uncertainties in the astrophysical modeling \kc{of...}.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

@@

%Universal relations
Unfortunately, current detectors are only sensitive enough to accurately measure the dominant tidal parameter, $\tilde{\Lambda}$, known as the chirp deformability\kc{That's the first time I hear this term!}.
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love universal relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction on the removal of one parameter from the model\kc{decrease in uncertainty in which parameter?}, and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of this relations leads to up to an order of magnitude improvement in parameter estimation \kc{of what?} through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis of~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" universal relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment , moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.

\kc{Say here that these relations are tuned to many EoS models and have intrinsic uncertainties. }
%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in~\cite{LIGO:posterior}, we repeat the analyses done in~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for both binary Love and I-Love-Q universal relations.
In addition, we consider hybrid star EoSs seen in~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into them.
Further, we estimate the viability of using improved universal relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in universal relations.

@@

%What we did
In this paper we attempt to improve binary Love and I-Love-Q universal relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love universal relations.
Finally, through a simple Fisher analysis, we consider the value in improving universal relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in universal relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
Additionally, we find that hybrid quark-hadron stars do obey the I-Love-Q and C-Love universal relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.
Unfortunately\kc{??}, due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the universal relations are far outweighed by the statistical errors accrued by extraction of high-order PN tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from parameter extraction for GW170817 as if it had been observed by future detectors $A \equiv ($ O2, aLIGO, a\texttt{++}, Voyager, ET, CE$)$\zc{cite these individually}\kc{a\texttt{++} should be A\texttt{+}}.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated interferometer $A$)\kc{How are we combining the uncertainty on $\tilde{\Lambda}$ when each event will have a different $\tilde{\Lambda}$?}.
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the universal relation systematics (indicated by the magenta horizontal line) for the Voyager telescope - indicating when improved binary Love universal relations will become viable\kc{viable? you mean necessary?}.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties $\sigma^A_{\text{GW170817}}$ (blue) in the extraction of $\tilde{\Lambda}$ from GW170817 (O2) as if observed with future interferometers aLIGO, a++, Voyager, CE, and ET-D as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$.
This is compared to the combined uncertainty $\sigma^A_N$ (red) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper (teal), central (orange), and lower (magenta) limits of expected binary NS merger detection rates for O2, aLIGO, a\texttt{++}, Voyager, CE, and ET-D over 1 observation year \kc{Instead of points, it might be clearer to show this with shaded regions from the upper to the lower rate.}.
It can be seen that the statistical errors in parameter extraction become comparable to the systematics from EoS variation in binary Love relations (purple shaded region) for Voyager and on.
Note there is no simulated populations of events under the O2 design sensitivity - due to there only being a single event during the observing period.
These results are further tabulated in Table~\ref{tab:variances}.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with a complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love universal relations, and considering how well hybrid star EoSs agree with the universal relations.
We next examine these improved universal relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations\kc{cite reviews by Baym and Oertel}, one method to study a wide range of physically realizable EoSs is to parameterize them in a model-independent\kc{how is this model-independent?} way.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior,Carney:2018sdv} sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations\kc{cite}.
We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.\zc{Carl, can you describe how you generated these? They aren't the \emph{actual} posteriors, are they?}\kc{Yeah they are}
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.

@@

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, important previous work by Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These relations, known as ``binary Love universal relations" allows one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) allows us to algebraically eliminate tidal parameters from the template parameter list, improving parameter estimation, and (ii) automatic measurement of second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love universal relations improved parameter extraction on $\tilde{\Lambda}$ by an order of magnitude.

Similar universal relations have been found to exist between individual NS observables: moment of inertia (I), tidal deformability (Love), quadrupole moment (Q), and compactness (C), known as the ``I-Love-Q" and ``C-Love" universal relations~\cite{Yagi:ILQ, Yagi:binLove}.
These have been found to be EoS insensitive by up to 1\% and 6.5\% respectively.
In this paper, we show improvement in both of these universal relations by restricting to EoSs sampled from GW170817s posterior on pressure as a function of density.
%%%%%%%%%%%%%%%%%%%%%%%%% Begin universal relations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Universal relations}\label{sec:universal}
In this section, we follow the analyses performed in~\cite{Yagi:binLove,Yagi:ILQ} with two new sets of 100 spectrally generated EoS: (i) posterior samples randomly selected from GW170817s 90\% posterior on on pressure as a function of density as shown in the left panel of Fig.~\ref{fig:eos}, and (ii) those unconstrained by any prior EoS information, as shown in the right panel of Fig.~\ref{fig:eos}.\kc{only one panel}

>>>>

\title{EoS-insensitive relations after GW170817}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Begin Abstract %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% Intro
The thermodynamic relationship between pressure and density (the equation of state, or EoS) of cold supranuclear matter, accessible in neutron stars, is critical to the study of such stars, and is one of the largest uncertainties in nuclear physics to date. 
% How to probe
The extraction of tidal deformabilities from the gravitational waveforms of binary neutron star merger events, such as GW170817, is a promising method of probing the nuclear structure.
% What was done before - Binary Love
Previous studies have shown that approximately EoS-insensitive ``binary Love universal relations" exist between symmetric and antisymmetric combinations of individual tidal deformabilities, which are difficult to be independently measured with second-generation gravitational wave interferometers.
% What was done before - I-Love-Q
Similarly, another set of EoS-insensitive relations exist between individual neutron star parameters: moment of inertia (I), tidal deformability (Love number), quadrupole moment (Q), and compactness (C) - known as the ``I-Love-Q" and ``C-Love" universal relations.
% What universal relations do
Such EoS-insensitive relations allow the elimination of some tidal parameters from the list of model parameters, thus breaking degeneracies and ultimately reducing the uncertainty of the dominant tidal effect in parameter estimation\kc{You mean $\tilde{\Lambda}$? In our paper with Carl and Aaron we showed that the binary Love relation doesn't affect the $\tilde{\Lambda}$ measurement much.}.
% What we will do
\kc{Add a sentence saying that even the more precise relations have some EoS variation which can cause a systematic error which can be marginalized over}
In this document, we explain how one can reduce equation of state variation for I-Love-Q, C-Love, and binary Love EoS-insensitive relations, which helps to reduce systematic errors on the tidal measurement of binary neutron star mergers.
However, we show that the systematic effects of these relations is very small until next-generation detectors. 
We achieve this  by restricting to only equations of state drawn from the 90\% posterior constraint on pressure as a function of density, as derived by the LIGO Collaboration.
% What we found
We find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with a mass ratio of 0.75, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim 50$\%.
% Conclude
We end by comparing systematic errors on tidal measurement due to the equation of state variation with statistical errors and comment on whether one can safely use these EoS-insensitive relations with future gravitational wave observations.
We conclude that the systematic errors injected from the use of improved EoS-insensitive relations become comparable to the statistical errors for future design upgrade Voyager.
\end{abstract}

@@

\section{Introduction}\label{sec:intro}
%Intro
The dependence of pressure on density for cold supranuclear matter found inside neutron stars (NSs) remains to be one of the largest mysteries in both nuclear physics and astrophysics.
This internal structure of such stars, known as the equation of state (EoS), is exceedingly important, as it is the determining factor for many NS observables - such as the mass, the radius, and many more.
Unfortunately, terrestrial experiments can only study the EoS up to the nuclear saturation density ($\rho_{\text{sat}} \approx 2.5 \times 10^{14} \text{ g/cm}^3$)~\cite{Li:HeavyIon,Tsang:SymmetryEnergy,Centelles:NeutronSkin,Li:CrossSections,Chen:SymEnergy}, making NSs ideal laboratories for constraining ultra dense nuclear matter\kc{heavy ion collisions supposedly go to higher densities, but they are T-dependent}.

%Constraining the EoS
Independent measurements of NS observables can be used to constrain the nuclear EoS.
For example, electromagnetic observations of the mass-radius relationship of NSs have been used to place limits on the NS EoS~\cite{guver,ozel-baym-guver,steiner-lattimer-brown,Lattimer2014,Ozel:2016oaf}.
However, these implications potentially suffer from large systematic errors due to astrophysical mismodeling uncertainties\kc{of...}.
Alternatively, the emission of gravitational waves (GWs) from binary NS merger events have proven to be a unique method of probing nuclear physics.
During the early inspiral portion of the merger, the orbital separation is large enough that the effect of companion tidal fields is negligible. 
As the separation decreases through GW emission, tidal forces are magnified and the NSs deform from sphericity, altering the orbital trajectories; a relic which is directly imprinted on the GW signal.
This deformation is characterized by the \textit{tidal deformability} $\Lambda$ as the linear response of the NSs quadrupole moment $Q_{ij}$ to the neighboring tidal field $\varepsilon_{ij}$~\cite{hinderer-love,Flanagan2008}.

@@

%Universal relations
Unfortunately, current detectors are only sensitive enough to accurately measure the dominant tidal parameter, $\tilde{\Lambda}$.
Previous work by Yagi and Yunes~\cite{Yagi:binLove} resolved this issue by finding approximately EoS-insensitive ``binary Love universal relations" between symmetric and anti-symmetric combinations of tidal deformabilities $\Lambda_{s,a}=\frac{1}{2}(\Lambda_1 \pm \Lambda_2)$.
This allows one to further break degeneracies between tidal parameters, resulting in both (i) a decrease in uncertainty upon parameter extraction of $\Lambda_s$ on the removal of parameter $\Lambda_a$ from the model, and (ii) the ability to measure the secondary tidal parameter upon measurement of the first~\cite{Katerina:residuals}.  
Use of these relations leads to up to an order of magnitude improvement in parameter estimation of $\Lambda_s$ through a simple Fisher analysis~\cite{Yagi:binLove}, also confirmed with the detailed analysis of~\cite{Katerina:residuals}.
Similar ``I-Love-Q" and ``C-Love" EoS-insensitive relations~\cite{Yagi:ILQ} have been found between individual NS observables: quadrupole moment , moment of inertia, tidal deformability, and compactness; similarly allowing the removal of tidal parameters from the model list, and allowing an automatic determination of one observable through the measurement of another.
However, these EoS-insensitive relations have been tuned to a collection of EoS models, and each have intrinsic uncertainties associated with their use.

%How we improved this
In this analysis, we aim to improve this important work by imposing the restrictions found on the EoS~\cite{LIGO:posterior} derived from the recent binary NS merger observation, GW170817~\cite{TheLIGOScientific:2017qsa}.
By restricting to only EoSs which lie within the the 90\% credible region on pressure as a function of density found in~\cite{LIGO:posterior}, we repeat the analyses done in~\cite{Yagi:binLove,Yagi:ILQ} and show an increase in universality for I-Love-Q, C-Love, and binary Love EoS-insensitive relations.
In addition, we consider hybrid star EoSs seen in~\cite{Paschalidis2018}, which experience strong first-order transitions from hadronic matter to quark matter.
These provide a departure from the nuclear matter relations considered here - providing valuable insight into them.
Further, we estimate the viability of using improved EoS-insensitive relations on future GW detections by performing simple Fisher analyses to approximate when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.

@@

%What we did
In this paper we attempt to improve I-Love-Q, C-Love, and binary Love EoS-insensitive relations by imposing constraints on the EoS found by GW170817~\cite{LIGO:posterior,TheLIGOScientific:2017qsa}.
First, we generate two large samples of spectral EoSs~\cite{Lindblom:2018rfr}, where we (i) impose the restriction that they must be drawn from GW170817s 90\% credible posterior in pressure as a function of density, and (ii) don't impose any restrictions for comparison purposes.
Next, we follow the important works of~\cite{Yagi:binLove} and~\cite{Yagi:ILQ} to show how the ``constrained" set of EoSs show improved universality (binary Love, I-Love-Q, and C-Love) from both the previous works, and the ``unconstrained" sample, shown by Fig.~\ref{fig:ILQ}.
In addition, we determine the degree to which hybrid star EoSs obey binary Love EoS-insensitive relations.
Finally, through a simple Fisher analysis, we consider the value in improving EoS-insensitive relations by approximating when the statistical errors from parameter extraction become comparable to the systematic errors due to EoS variation in EoS-insensitive relations.

%Final Results
Upon use of the constrained set of EoSs, we find an improvement in binary Love universality by a factor of $\sim 59$\% for stars with mass ratios of $0.75$, in C-Love universality by a factor of $\sim 67\%$, and in I-Love-Q universality by factors of $\sim50$\%, tabulated in Tab.~\ref{tab:maxVar}.
Additionally, we find that hybrid quark-hadron stars do obey the I-Love-Q and C-Love EoS-insensitive relations, with sightly increased EoS variability.
We also find that binaries consisting of one massive hybrid star, and one small-mass hadronic star do not agree with derived binary Love relations, due to the large separation in $\tilde{\Lambda}$ between the constituent stars.
Due to the current limitations in detector sensitivity, the systematic errors arising from EoS variation in the EoS-insensitive relations are far outweighed by the statistical errors accrued by extraction of high-order PN tidal parameters from GW waveforms.
Fig.~\ref{fig:stackedFisher} compiles the results from Fisher analyses approximating the uncertainties accrued from parameter extraction for GW170817 as if it had been observed by future detectors $A \equiv ($ O2, aLIGO, A\texttt{+}, Voyager, ET, CE$)$\zc{cite these individually}.
Here, $\sigma^A_{\text{GW170817}}$ and $\rho^A_{\text{GW170817}}$ correspond to the statistical error and signal-to-noise-ratio found from observing GW170817 on detector $A$, and $\sigma^A_N$ represents the combined uncertainty after the detection of $N_A$ events (corresponding to the binary NS merger rate associated interferometer $A$).
These values are further tabulated in Table~\ref{tab:variances}.
It can be seen from this figure that the statistical errors from parameter extraction become comparable the EoS-insensitive relation systematics (indicated by the magenta horizontal line) for the Voyager telescope - indicating when improved binary Love EoS-insensitive relations will become necessary.
\begin{figure}
\begin{center} 
\includegraphics[width=\columnwidth]{stackedFisher.eps}
\end{center}
\caption{
Estimated statistical uncertainties $\sigma^A_{\text{GW170817}}$ (blue circle) in the extraction of $\tilde{\Lambda}$ from GW170817 (O2) as if observed with future interferometers aLIGO, A\texttt{+}, Voyager, CE, and ET-D as a function of the signal-to-noise-ratio $\rho^A_{\text{GW170817}}$.
This is compared to the combined uncertainty $\sigma^A_N$ (turqoise shaded region) from the observation of multiple events from a simulated population of size $N_A$ corresponding to the upper and lower limits of expected binary NS merger detection rates for O2, aLIGO, A\texttt{+}, Voyager, CE, and ET-D over 1 observation year.
It can be seen that the statistical errors in parameter extraction become comparable to the systematics from EoS variation in binary Love relations (purple shaded region) for Voyager and on.
Note there is no simulated populations of events under the O2 design sensitivity - due to there only being a single event during the observing period.
These results are further tabulated in Table~\ref{tab:variances}.
}
\label{fig:stackedFisher}
\end{figure}

%outline
The organization of this paper is outlined below.
We begin with a complementary background and theory material in Sec.~\ref{sec:theory}.
We continue in Sec.~\ref{sec:universal} by finding new and improved binary Love, I-Love-Q, and C-Love EoS-insensitive relations, and considering how well hybrid star EoSs agree with the EoS-insensitive relations.
We next examine these improved EoS-insensitive relations and question whether or not they are useful for future interferometers in Sec.~\ref{sec:observations}.
We conclude in Sec.~\ref{sec:conclusion} by discussing our results and mentioning avenues of future work.
Throughout this paper, we have adopted geometric units of $G=c=1$, unless otherwise stated.

@@

\subsection{Spectral representations of neutron star equations of state}\label{sec:eos}
The structure of a NS and its tidal interactions in a binary system rely heavily on the underlying state function describing the relationship between pressure ($p$) and energy density ($\epsilon$), or equation of state, of nuclear matter.
Given that all currently proposed EoSs utilize certain approximations\kc{cite reviews by Baym and Oertel}, one method to study a wide range of physically realizable EoSs is to parameterize them in a model-independent way\kc{how is this model-independant?}.
Spectral representations~\cite{Lindblom:2010bb,Lindblom:2012zi,Lindblom:2013kra,Lindblom:2018rfr,Abbott:2018exr} parameterize EoSs by performing spectral expansions on the adiabatic index $\Gamma(p)$\footnote{Another way of parameterizing EoSs is the piecewise polytropic formulation~\cite{Read2009,Lackey:2014fwa,Carney:2018sdv}.}:
\begin{equation}
\Gamma(x) = \exp{\sum_k\gamma_k x^k},
\end{equation}
where $x \equiv \log{p/p_0}$ for minimum pressure $p_0$.
The equation of state is then determined by an integration of the differential equation:
\begin{equation}
\frac{d \epsilon(p)}{dp}=\frac{\epsilon(p)+p}{p \Gamma(p)}.
\end{equation}
Using this formalism, any valid EoS can be approximated through choice of 4 or more spectral coefficients $\gamma_k$, tabulated for several common EoSs in Table 1 of~\cite{Lindblom:2018rfr}.

In the current analysis, we consider EoSs which have been constrained by the recent binary neutron star merger event, GW170817.
Important recent work~\cite{LIGO:posterior,Carney:2018sdv} sampled EoS parameters in order to derive a marginalized posterior on the pressure as a function of mass density, as seen in Fig. 2 of~\cite{LIGO:posterior}.
The spectral coefficients were sampled within the ranges: $\gamma_0 \in \lbrack 0.2,2 \rbrack$, $\gamma_1 \in \lbrack -1.6,1.7 \rbrack$, $\gamma_2 \in \lbrack -0.6,0.6 \rbrack$, $\gamma_0 \in \lbrack -0.02,0.02 \rbrack$, and further restricted the adiabatic index to be $\Gamma \in \lbrack 0.6,4.5 \rbrack$, ensuring the parameterization exposed a wide range of viable EoSs~\cite{Lindblom:parameters}.
Additional constraints imposed upon the generated EoSs were as follows~\cite{LIGO:posterior}: (i) causality within 10\%, and (ii) EoS priors must support NS masses up to $1.97 \text{ M}_{\odot}$, consistent with astrophysical observations\kc{cite}.
We utilize a random set of 100 of the above posterior samples defined by GW170817, furthermore referenced as the ``constrained EoSs", as shown in Fig.~\ref{fig:eos}.
Following~\cite{Read2009}, the parameterized high-density core EoSs generated above are matched to the low-density crust EoS Sly~\cite{Douchin:2001sv} at about half of the nuclear saturation density, $\rho_{\text{stitch}}=1.3 \times 10^{14} \text{ g/cm}^3$.
For comparison, we also include a second sample of 100 ``unconstrained" EoSs, randomly sampled in the pressure-density plane to be unrestricted by GW170817s posterior.

@@

\subsection{EoS insensitive relations}\label{sec:eosInsensitive}
Current gravitational wave interferometry is not yet sensitive enough to accurately extract both tidal parameters $\tilde{\Lambda}$ and $\delta\tilde{\Lambda}$.
In a search to remedy this, important previous work by Yagi and Yunes~\cite{Yagi:binLove} found that symmetric and antisymmetric combinations of tidal deformabilities:
\begin{equation}
\Lambda_s \equiv \frac{\Lambda_1 + \Lambda_2}{2}, \hspace{6mm} \Lambda_a \equiv \frac{\Lambda_1 - \Lambda_2}{2},
\end{equation}
display EoS-insensitive properties to a high degree, showing EoS variations up to 20\% for binaries with masses less than $1.7 \text{ M}_{\odot}$, for a representative set of 11 EoSs. 
These relations, known as ``binary Love universal relations" allows one to break degeneracies between coupled tidal parameters.
This is important for two reasons: (i) allows us to algebraically eliminate tidal parameters from the template parameter list, improving parameter estimation, and (ii) automatic measurement of second tidal parameter given observation of the first.
It was shown through a simple Fisher analysis that the use of binary Love EoS-insensitive relations improved parameter extraction on $\tilde{\Lambda}$ by an order of magnitude.

Similar EoS-insensitive relations have been found to exist between individual NS observables: moment of inertia (I), tidal deformability (Love), quadrupole moment (Q), and compactness (C), known as the ``I-Love-Q" and ``C-Love" EoS-insensitive relations~\cite{Yagi:ILQ, Yagi:binLove}.
These have been found to be EoS insensitive by up to 1\% and 6.5\% respectively.
In this paper, we show improvement in both of these EoS-insensitive relations by restricting to EoSs sampled from GW170817s posterior on pressure as a function of density.
%%%%%%%%%%%%%%%%%%%%%%%%% Begin universal relations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{EoS-insensitive relations}\label{sec:universal}
In this section, we follow the analyses performed in~\cite{Yagi:binLove,Yagi:ILQ} with two new sets of 100 spectrally generated EoS: (i) posterior samples randomly selected from GW170817s 90\% posterior on on pressure as a function of density as shown by the solid green curves in Fig.~\ref{fig:eos}, and (ii) those unconstrained by any prior EoS information, as shown by the dotted maroon curves in Fig.~\ref{fig:eos}.

\subsection{I-Love-Q relations}\label{sec:ilq}
Here we present our results on I-Love-Q universality as compared to Fig. 1 of Yagi and Yunes~\cite{Yagi:ILQ}.
In particular, we consider two distinct classes of NSs: nuclear matter EoSs and hybrid quark-hadron star EoSs as described in Sec.~\ref{sec:theory}.
We begin in Sec.~\ref{sec:ilq-nuc} by fitting the new I-Love-Q EoS-insensitive relations using the constrained set of EoSs.
This is followed in Sec.~\ref{sec:ilq-hyb} by an analysis and discussion into how well hybrid stars agree with the improved binary Love relations.
